---
title: "NYC Schools Perceptions"
author: "Thais Lovisi"
date: "08/12/2022"
output: html_notebook
---

## Introduction 

With demographic and test score data from the New York City Department of Education, this project aims to explore the data using visualization techniques and calculated correlations to understand how schools' demographics affect SAT scores, an indicator of schools' academic performance are related.

This dataset also has some data about: Responses to surveys designed to gauge parent, student, and teacher perceptions of the quality of New York City schools. 

In this project, we are targeting to answer the following questions:

  Do student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics?

  Do students, teachers, and parents have similar perceptions of NYC school quality?

## Step 0 - Setup  
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library (tidyverse)
library(dplyr)
library("readxl")
library(plyr)
library(kableExtra)

```
###Loading files:
    Files information:
      **Survey Data Dictionary.xls**: contains metadata useful for decision making about how to clean and prepare the survey data for analysis.
      **masterfile11_gened_final.xlsx**: contain survey data for "general education" schools — those that do not specifically serve populations with special needs.
      **masterfile11_d75_final.xlsx**: contain survey data for District 75 schools, which provide special education support for children with special needs such as                learning or physical disabilities.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
survey_special_school <- read_excel("masterfile11_d75_final.xlsx")
general_education <- read_excel ("masterfile11_gened_final.xlsx")
survey <- read_excel ("Survey Data Dictionary.xls")
```
### Dataset Familiarisation

#### For **survey** file

      Identifying data type
      
```{r}
for (c in colnames(survey)){
  print(typeof(survey[[c]]))
}

```
        Column's Names
        
```{r}
colnames(survey)
survey <- survey[c(-1,-2,-3),]
```
        Identifying unique values for each column
        
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
head(survey)
tail(survey)
```       
       Renaming columns
    
```{r}
colnames(survey) <- c("Survey_Field", "Description")
```

#### For **general_education** file
       Initial dimensions
    
```{r echo=TRUE}
dim(general_education)
ncol(general_education)
nrow(general_education)
```     
      Column's Names
        
For general_education colnames displayed not the variable names hence the survey variables are displayed at row 2 not at the first row. Thus is needed to remove row 1 and the header.

```{r}
general_education <- general_education[-1,]
colnames(general_education) <- general_education[2,]
```

 
        
#### Cleaning Unessecery Columns 

Taking in to account that our vlook up has the transposed value in correct order we can identify the variables labled and clean the ones that we cannot identify.

Taking a look at the Data Dctionary , in survey we see that have some variables of interest not listed by the year 2011(_11) but detailed to 2010(_10), in this way
the vlookup didn`t added their names. 

```{r}
#if I wanted the list of columns too keep at the dataset I would do:
C_Kept <- for (c in 2){
                for (i in 1:nrow(Look_up_step1)){
                    if (is.na(Look_up_step1[i,c])) {
                      next #if doesnt work with other coders go change break by next
                    }else{
                      print (c(i))
                      next
                    }
                }       
}


```
So lets again take a look on our variables of interest :

      dbn 
      schoolname 
      schooltype 	
      Safety and Respect score based on parent responses (saf_p_10)
      Communication score based on parent responses (com_p_10)
      Engagement score based on parent responses (eng_p_10)
      Academic expectations score based on parent responses(aca_p_10)
      Safety and Respect score based on teacher responses (saf_t_10) 
      Communication score based on teacher responses (com_t_10)
      Engagement score based on teacher responses (eng_t_10)
      Academic expectations score based on teacher responses (aca_t_10)
      Safety and Respect score based on student responses (saf_s_10)
      Communication score based on student responses (com_s_10)
      Engagement score based on student responses (eng_s_10)
      Academic expectations score based on student responses(aca_s_10)
      Safety and Respect total score (saf_tot_10)
      Communication total score (com_tot_10)
      Engagement total score (eng_tot_10)
      Academic Expectations total score (aca_tot_10)

so lets select those variables filtered by high school`s to create a new_general_education data set

```{r}
new_general_education <- general_education %>% select(dbn,schoolname, saf_p_11,com_p_11,eng_p_11,aca_p_11,saf_t_11,com_t_11,eng_t_11,aca_t_11,saf_s_11,com_s_11,	
eng_s_11,aca_s_11,saf_tot_11,com_tot_11,eng_tot_11,aca_tot_11)
```
   
   We need filter as well by schooltype, how can we identify the school type on this data set?
          Lets create a new column where if the school name contains "High" or "High School", the column will have the input H, otherwise NA.

```{r}
High_School <- function(schoolname){
  hschool_positive = case_when(
    str_detect(schoolname, "High") ~ TRUE,
    str_detect(schoolname, "High School") ~ TRUE,
  TRUE ~ FALSE
    )
}

new_general_education <- new_general_education[-1,]

new_general_education <- new_general_education %>% #Rewriting it in a new colunm called positive_Review
  mutate(
   stype = unlist(map(schoolname, High_School))
  )

new_general_education <- new_general_education %>%
        filter(stype == TRUE)
new_general_education <- new_general_education[,-19]

```

After the initial data cleaning we have 255 observations for 19 variables at the data set that refers to the General Education High Schools (new_general_education).


#### For **survey_special_school** file

        Identifying unique values for each column
Initial dimensions
    
```{r echo=TRUE}
dim(survey_special_school)
ncol(survey_special_school)
nrow(survey_special_school)
```     
      Column's Names
        
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
colnames(survey_special_school)
```

There is some column code identified on Survey Data Dictionary? 
      Yes, and are the same as before.
      
```{r}
new_special_education <- survey_special_school[-1,]


colnames(new_special_education) <- new_special_education[1,]

new_special_education <- new_special_education %>% 
  select(dbn,schoolname,saf_p_11,com_p_11,eng_p_11,aca_p_11,saf_t_11,com_t_11,eng_t_11,aca_t_11,saf_s_11,com_s_11,	
  eng_s_11,aca_s_11,saf_tot_11,com_tot_11,eng_tot_11,aca_tot_11) 


```


#### Combining survey data

Combine `new_general_education` and `new_special_education` data frames.
Since bind_rows(), like other dplyr functions, is designed for manipulating dataframes, it can be used in situations where dataframes have different numbers of variable columns. Using bind_rows() here, the output dataframe will contain a column if that column appears in any of the input dataframes,I'm using bind_rows() to combine "survey_data_gen and survey_data_dis into a single dataframe.

```{r}
Survey_Total <- new_general_education %>% 
  bind_rows(new_special_education)
```

The tables used in the next steps will be **Survey_Total**, **new__special_education** and **new_general_education**. Survey_Total is the table that contains merged data for both types of schools, with 18 cols and 312 rows.


### Step 1 - Handling missing Data

#### For Survey_total 

We have found NA data on the following columns :

```{r}
for (c in colnames(Survey_Total)){
  for (i in 1:nrow(Survey_Total)){
    if (is.na(Survey_Total[i,c])){
      print (c)
      break
    }else{
      next
    }
  }
}
``` 

The method of handling the missing data will be complete removal. We know by the data description that only students in grades 6-12 participated on the survey. As all missing data refers to variables related to student perception is possible that filter on variable will impact on the others. Thus in this step we will:
        Filter all in saf_s_11 that is not NA
        Use loop for to check if there is a NA value remaining.

```{r}
Survey_Total = Survey_Total%>%
  filter(!is.na(`saf_s_11`))
for (c in colnames(Survey_Total)){
  for (i in 1:nrow(Survey_Total)){
    if (is.na(Survey_Total[i,c])){
      print (c)
      break
    }else{
      next
    }
  }
}
``` 

Is not needed handle any additional NA. There is 305 rows and 18 cols at final table.

```{r}
a <- colnames(Survey_Total)
str_replace(a,"_10","_11")
colnames(Survey_Total) <- a
b <- t(survey[1])
b<- str_replace(b,"_10","_11")
survey[[1]] <- str_replace(survey[[1]],"location" , "schoolname")
survey[[1]] <- (b)
```
 There is some column code identified on Survey Data Dictionary?
```{r}
S_Fields <- (a) # Gets the first row and transform it on column
Vlook_up <-  as.data.frame(S_Fields,2)
colnames(Vlook_up) <- c("Survey_Field")
colnames(survey) <- c("Survey_Field","Description")
Look_up_step1 <- join(Vlook_up, survey, by = "Survey_Field")

# Add description to the table
Survey_Total_named <- rbind(Look_up_step1[[2]],Survey_Total)
```

Note that only 16 columns from dataset general_education were identified on the dictionary. 


Other way to do the vlook_up would be by using merge:
        **E.g.:** base1 <- (merge(survey, Vlook_up, by = 'Survey_Field'))

##Step 2 - Looking for Correlations 

Finding the correlation coefficient showing the parent, teacher and student perceptions of the following factors and how they affect average school SAT scores(an indicator of academic performance):

1. Safety
2. Engagement
3. Communication
4. Academics

#### With Scatter Plots

We need first create a Correlation Matrix(Correlation_Matrix) which is used to investigate the dependence between multiple variables at the same time. The result of a correlation matrix is a table containing the correlation coefficients between each variable and the others.

Easy way to apply a corr. matrix is doing Correlation_Matrix <- cor(my_data). 


```{r}
Conv_df <- as.data.frame(Survey_Total) # convert to data frame
i <- c(3:18)

Survey_Total[,i] <- apply(Survey_Total[,i],2, function(x) as.numeric(x)) # convert to numeric




```
Note that some NA values were added, it happened because when we merged the 2 data sets to generate the Survey_Total we bind together the header, lets pop this value out:

```{r}
for (c in colnames(Survey_Total)){
  for (i in 1:nrow(Survey_Total)){
    if (is.na(Survey_Total[i,c])){
      print (c(c,i))
      break
    }else{
      next
    }
  }
}

Survey_Total <- Survey_Total[-251,]

```

Now lets create our Correlation matrix

```{r echo=TRUE}
Correlation_matrix <- Survey_Total %>% select(saf_p_11:aca_tot_11) %>% cor(use = "pairwise.complete.obs")

# corr_mat_tib <- Correlation_matrix %>% as_tibble(rownames = "variable") / if i wanted convert in tibble
corr_mat_tib <- round(Correlation_matrix,3) #round values

#visualizing table

corr_mat_tib %>% 
  kable(bootstrap_options = c("striped", "condensed"),
        full_width = FALSE,
                font_size = 15,
        )%>% row_spec(0, angle = -15, color = "blue", font_size = 13)
```

"The closer correlation coefficient is to zero the weaker the relationship", with this I can say that there is a weak positive relationship between ...

